---
title: "Project2"
author: "LeiSun"
date: "2020/4/29"
output: html_document
---

```{r setup, include=FALSE}
setwd('F:\\Courses\\HKUST_Semester_Spring\\MAFS5210\\project2')
library(MASS)
library(data.table)
```

## Step1: 
Prepare the factor return data
```{r}

load('F:\\Courses\\HKUST_Semester_Spring\\MAFS5210\\project2\\data_for_R\\Rtn_fac.rda')
setorder(Rtn_fac,Date)
head(Rtn_fac)
dim(Rtn_fac)
```
```{r}
#write.csv(Rtn_fac,file = "F:\\Courses\\HKUST_Semester_Spring\\MAFS5210\\project2\\data_for_R\\Rtn_fac.csv",row.names = T)
```


##Step2: 
Factor orthogonalization & Factor covariance matrix F
```{r}
.rmRegress<-function(old.factors, new.factors){
  Givn=ginv(old.factors)
  fit3=Givn%*%new.factors
  
  resids=new.factors-old.factors%*%fit3
  factors=cbind(old.factors,resids)
  
  list(factors=factors,betas=fit3)
}

BM='benchmark'
Sct = c('金融行业','房地产', '综合行业', '电力行业', '机械行业', '有色金属', '农林牧渔', '电子信息', '酒店旅游', '家电行业', '传媒娱乐', '汽车制造', '发电设备', '水泥行业', '商业百货', '电子器件', '生物制药', '农药化肥', '酿酒行业', '环保行业', '化工行业', '钢铁行业', '化纤行业', '煤炭行业', '飞机制造', '建筑建材', '食品行业', '次新股', 'NaN', '仪器仪表', '服装鞋类', '家具行业', '交通运输', '其它行业', '印刷包装', '船舶制造', '医疗器械', '石油行业', '物资外贸', '塑料制品', '玻璃行业', '公路桥梁', '电器行业', '开发区', '供水供气')
Style=c('MOM','vol_volatility_1m','volatility_1m','pe','circ_mv','dv_ratio')
Macro=c('shibor_1m', 'shibor_6m', 'libor_1m', 'libor_6m','shibor_lpr_1y')

tempF=.rmRegress(old.factors = as.matrix(cbind(rep(1,60),Rtn_fac[,BM,with=F])),new.factors=as.matrix(Rtn_fac[,Sct,with=F]))
tempF=.rmRegress(old.factors = as.matrix(tempF$factors),new.factors = as.matrix(Rtn_fac[,Style,with=F]))
tempF=.rmRegress(old.factors = as.matrix(tempF$factors),new.factors = as.matrix(Rtn_fac[,Macro,with=F]))
Rtn_fac_orth=copy(tempF$factors[,-1])
head(Rtn_fac_orth)
F_matrix=var(Rtn_fac_orth)

```
```{r}
#write.csv(Rtn_fac_orth,file = "F:\\Courses\\HKUST_Semester_Spring\\MAFS5210\\project2\\data_for_R\\Rtn_fac_orth.csv",row.names = T)

```

```{r}
write.csv(F_matrix,file = "F:\\Courses\\HKUST_Semester_Spring\\MAFS5210\\project2\\data_for_R\\F_matrix.csv",row.names = T)
```


##Step3: 
Prepare the Stock return data
```{r}
Rtn_stk=read.csv("F:\\Courses\\HKUST_Semester_Spring\\MAFS5210\\project2\\data_for_R\\Rtn_stk.csv",
                 header = TRUE, fileEncoding  =  "UTF-8")
Rtn_stk=data.table(Rtn_stk)
Rtn_stk[,Date:=Rtn_stk$trade_date]
Rtn_stk=Rtn_stk[,Ticker:=substring(Rtn_stk$Ticker,1,6)]
Rtn_stk=Rtn_stk[,c('Date','Ticker','Sct','Return')]
Rtn_stk[,MTH:=paste0(year(Date),month(Date))]
#load('Sector.rda')
#Sector[,Ticker:=substring(Ticker,1,6)]
#Rtn_stk=merge(Rtn_stk,Sector,by='Ticker',all.x=TRUE)
#Rtn_stk[Rtn_stk$Sct=='NaN',Sct:='Others']
Rtn_stk
```


##Step4: 
Impute the missing stock returns
```{r}
load('F:\\Courses\\HKUST_Semester_Spring\\MAFS5210\\project2\\data_for_R\\Rtn_fac.rda')
Rtn_fac[,Others:=Rtn_fac[,'NaN']]
Rtn_fac=Rtn_fac[,-2]
Rtn_fac = melt(Rtn_fac,id.vars = 'Date')
Rtn_fac[,MTH:=paste0(year(Date),month(Date))]
remove=c('2019-02-28')
Rtn_stk=Rtn_stk[!Date %in% remove,]
setkey(Rtn_stk,Sct,MTH)
setkey(Rtn_fac,variable,MTH)
Rtn_stk=Rtn_fac[Rtn_stk]
Rtn_stk[is.na(Return),Return:=value]
Rtn_stk
Rtn_stk=dcast(Rtn_stk,Date~Ticker,value.var = 'Return')

```

##Step5: 
Calculate stock beta against factors: B, where $r_t=Bf_t+\theta_t$
```{r}
reg = .rmRegress(old.factors = as.matrix(Rtn_fac_orth),new.factors = as.matrix(Rtn_stk[,-1,with=F]))
B=t(reg$betas)
rownames(B)=paste(rownames(B),'CH',sep='.')
colnames(B)=colnames(Rtn_fac_orth)
head(B)

```

##Step6:
Caculate idiosyncratic variance matrix: D
```{r}
K=1+length(Sct)+length(Style)+length(Macro)
D=diag(as.numeric(apply(reg$factors[,-(1:K)],2, function(x) sum(x^2)/(60-K))))
head(D)
```
##Step7:
Calculate factor loading: $u=B^Tw$
```{r}
Weight=read.csv('F:\\Courses\\HKUST_Semester_Spring\\MAFS5210\\project2\\data_for_R\\Weight.csv')
Weight=data.table(Weight)
Weight=Weight[,c('symbol','Wgt_Port','Wgt_BM','Wgt_Act')]
w=cbind(Weight$Wgt_BM,Weight$Wgt_Port,Weight$Wgt_Act)
B=as.data.frame(B)
B[is.na(B)] <- 0
B=as.matrix(B)
u=t(B)%*%w
u
  
```


```{r}
#write.csv(u,file = "F:\\Courses\\HKUST_Semester_Spring\\MAFS5210\\project2\\data_for_R\\data_loading.csv",row.names = T)
```

##Step8:
Total risk attribution
1)
```{r}
Weight=read.csv('F:\\Courses\\HKUST_Semester_Spring\\MAFS5210\\project2\\data_for_R\\Weight.csv')
Weight=data.table(Weight)
Weight=Weight[,c('symbol','Wgt_Port','Wgt_BM','Wgt_Act')]
D=as.data.frame(D)
D[is.na(D)] <- 0
D=as.matrix(D)
w=as.matrix(Weight$Wgt_Port)
TV_Sys=as.numeric(t(w)%*%B%*%F_matrix%*%t(B)%*%w)*12
TV_Spe=t(w)%*%D%*%w*12
TR=as.numeric(sqrt(TV_Sys+TV_Spe))
TV_Sys
TV_Spe
TR


```
2)
```{r}
TV_BM=as.numeric(t(w)%*%B[,BM]%*%F_matrix[BM,BM]%*%t(B[,BM])%*%w)*12
TV_Sct=as.numeric(t(w)%*%B[,Sct]%*%F_matrix[Sct,Sct]%*%t(B[,Sct])%*%w)*12
TV_Sty=as.numeric(t(w)%*%B[,Style]%*%F_matrix[Style,Style]%*%t(B[,Style])%*%w)*12
TV_Mac=as.numeric(t(w)%*%B[,Macro]%*%F_matrix[Macro,Macro]%*%t(B[,Macro])%*%w)*12
TV_BM
TV_Sct
TV_Sty
TV_Mac
TV_BM+TV_Sct+TV_Sty+TV_Mac-TV_Sys
```
3)
```{r}
Factors=c(BM,Sct,Style,Macro)
TV=sapply(Factors,function(fac){
  f=matrix(0,ncol=ncol(F_matrix),nrow = nrow(F_matrix))
  f[which(fac==Factors),]=F_matrix[which(fac==Factors),]
  as.numeric(t(w)%*%B%*%f%*%t(B)%*%w)*12
})
TV
sum(TV[Sct])-TV_Sct
sum(TV[Style])-TV_Sty
sum(TV[Macro])-TV_Mac



```
```{r}
#write.csv(TV,file = "F:\\Courses\\HKUST_Semester_Spring\\MAFS5210\\project2\\data_for_R\\Total_risk_attribution.csv",row.names = T)
```

4)
```{r}
V=(B%*%F_matrix%*%t(B)+D)*12
MCTR=V%*%w/TR
head(MCTR)
sum(w*MCTR)-TR
```
```{r}
#write.csv(MCTR,file = "F:\\Courses\\HKUST_Semester_Spring\\MAFS5210\\project2\\data_for_R\\MCTR.csv",row.names = T)
```


##Step9:
Tracking error attribution
1)
```{r}
w=as.matrix(Weight$Wgt_Act)
ActV_Sys=as.numeric(t(w)%*%B%*%F_matrix%*%t(B)%*%w)*12
ActV_Spe=t(w)%*%D%*%w*12
TE=as.numeric(sqrt(ActV_Sys+ActV_Spe))
ActV_Sys
ActV_Spe
TE


```
2)
```{r}
ActV_BM=as.numeric(t(w)%*%B[,BM]%*%F_matrix[BM,BM]%*%t(B[,BM])%*%w)*12
ActV_Sct=as.numeric(t(w)%*%B[,Sct]%*%F_matrix[Sct,Sct]%*%t(B[,Sct])%*%w)*12
ActV_Sty=as.numeric(t(w)%*%B[,Style]%*%F_matrix[Style,Style]%*%t(B[,Style])%*%w)*12
ActV_Mac=as.numeric(t(w)%*%B[,Macro]%*%F_matrix[Macro,Macro]%*%t(B[,Macro])%*%w)*12
ActV_BM
ActV_Sct
ActV_Sty
ActV_Mac
ActV_BM+ActV_Sct+ActV_Sty+ActV_Mac-ActV_Sys
```
3)
```{r}
Factors=c(BM,Sct,Style,Macro)
ActV=sapply(Factors,function(fac){
  f=matrix(0,ncol=ncol(F_matrix),nrow = nrow(F_matrix))
  f[which(fac==Factors),]=F_matrix[which(fac==Factors),]
  as.numeric(t(w)%*%B%*%f%*%t(B)%*%w)*12
})
ActV
sum(ActV[Sct])-ActV_Sct
sum(ActV[Style])-ActV_Sty
sum(ActV[Macro])-ActV_Mac
```
```{r}
#write.csv(ActV,file = "F:\\Courses\\HKUST_Semester_Spring\\MAFS5210\\project2\\data_for_R\\Tracking_error_attribution.csv",row.names = T)
```

4)
```{r}
MCAR=V%*%w/TE
head(MCAR)
sum(w*MCAR)-TE
```
```{r}
#write.csv(MCAR,file = "F:\\Courses\\HKUST_Semester_Spring\\MAFS5210\\project2\\data_for_R\\MATR.csv",row.names = T)
```

## PCA
```{r}
Rtn_stk2=read.csv('F:\\Courses\\HKUST_Semester_Spring\\MAFS5210\\project2\\data_for_R\\Rtn_stk2.csv')
Rtn_stk2=data.table(Rtn_stk2)

Rtn_stk2[,length(unique(Ticker))]
Rtn_stk2=dcast(Rtn_stk2,Date~Ticker,value.var = 'Return_m')

Rtn_stk2
```

```{r}
Rtn_stk2[is.na(Rtn_stk2)] <- 0
V=cov(Rtn_stk2[,-1,with=F])
E=eigen(V)
sqrt(E$values)

```
```{r}
aa=sqrt(E$values)
#write.csv(aa,file = "F:\\Courses\\HKUST_Semester_Spring\\MAFS5210\\project2\\data_for_R\\PCA.csv",row.names = T)

```







##Step10:
Lower the tracking error
reduce the exposure in sectors '电子器件' & '传媒娱乐'
```{r}
Weight=read.csv('F:\\Courses\\HKUST_Semester_Spring\\MAFS5210\\project2\\data_for_R\\Weight.csv')
Weight=data.table(Weight)
Weight[,Ticker:=Weight[,symbol]]
Weight=Weight[,c('Ticker','Wgt_Port','Wgt_BM','Wgt_Act')]
Weight[,Ticker:=substring(Ticker,1,6)]
Sector=read.csv('F:\\Courses\\HKUST_Semester_Spring\\MAFS5210\\project2\\data_for_R\\Sector.csv',fileEncoding = 'UTF-8')
Sector=data.table(Sector)
Sector[,Ticker:=substring(Ticker,1,6)]
Weight=merge(Weight,Sector,by='Ticker',all.x=TRUE)
Weight[,Wgt_Port_adj:=Wgt_Port]

Adj1=Weight[Sct=='电子器件', .(sum(Wgt_Port),sum(Wgt_BM))]

Weight[Sct=='电子器件',Wgt_Port_adj:=Wgt_Port*(Adj1$V2+0.1)/Adj1$V1]

Adj2=Weight[Sct=='传媒娱乐', .(sum(Wgt_Port),sum(Wgt_BM))]
Weight[Sct=='传媒娱乐',Wgt_Port_adj:=Wgt_Port*(Adj2$V2+0.1)/Adj2$V1]

Weight[(Sct!='电子器件')&(Sct!='传媒娱乐'),Wgt_Port_adj:=Wgt_Port_adj/sum(Wgt_Port_adj)*Weight[(Sct=='电子器件')|(Sct=='传媒娱乐'),1-sum(Wgt_Port_adj)]]

Weight[,sum(Wgt_Port_adj-Wgt_BM),by=Sct]

```
```{r}
Weight[,Wgt_Act_adj:=Wgt_Port_adj-Wgt_BM]
w=as.matrix(Weight$Wgt_Act_adj)
ActV_Sys=as.numeric(t(w)%*%B%*%F_matrix%*%t(B)%*%w)*12
ActV_Spe=t(w)%*%D%*%w*12
ActV_Sys
ActV_Spe
TE=as.numeric(sqrt(ActV_Sys+ActV_Spe))
TE
```
##Step11:
Increase stock specific risk

1)Identify the source of the stock specific risk
```{r}
StockSpec=apply(reg$factors[,-(1:K)],2,function(x) sum(x^2)/(60-K))
StockSpec=data.table(Ticker=paste0(names(StockSpec),'.SH'),SSVar=StockSpec)
StockSpec[is.na(StockSpec)] <- 0
Weight=read.csv('F:\\Courses\\HKUST_Semester_Spring\\MAFS5210\\project2\\data_for_R\\Weight.csv')
Weight=data.table(Weight)
Weight=Weight[,c('symbol','Wgt_Port','Wgt_BM','Wgt_Act')]
setkey(Weight,symbol)
setkey(StockSpec,Ticker)
Weight=StockSpec[Weight]
Weight[Wgt_Port>0,Rank:=frankv(SSVar,order=-1)]
setorder(Weight,Rank,na.last = TRUE)
Weight
```
```{r}
#write.csv(Weight,file = "F:\\Courses\\HKUST_Semester_Spring\\MAFS5210\\project2\\data_for_R\\SSVar.csv",row.names = T)
```

2)Increase the exposure to those stocks
```{r}
Weight[,Wgt_Port_adj:=Wgt_Port]
Weight[Rank<=10,Wgt_Port_adj:=Wgt_Port_adj*2]
Weight[Rank>10,Wgt_Port_adj:=Wgt_Port_adj/sum(Wgt_Port_adj)*Weight[Rank<=10,1-sum(Wgt_Port_adj)]]

Weight[,Wgt_Act_adj:=Wgt_Port_adj-Wgt_BM]
w=as.matrix(Weight$Wgt_Act_adj)
ActV_Sys=as.numeric(t(w)%*%B%*%F_matrix%*%t(B)%*%w)*12
ActV_Spe=t(w)%*%D%*%w*12
ActV_Sys
ActV_Spe
ActV_Spe/(ActV_Spe+ActV_Sys)
TE=as.numeric(sqrt(ActV_Sys+ActV_Spe))
TE
```
##Step12:
Minimum Tracking error portfolio
objectie function & gradient of it, equality constraint function & gradient of it, 
```{r}
ObjFun=function(w){
  w1=as.matrix(w-Weight$Wgt_BM)
  as.numeric(t(w1)%*%B%*%F_matrix%*%t(B)%*%w1+t(w1)%*%D%*%w1)
}

ObjFunGradient = function(w){
  w1=as.matrix(w-Weight$Wgt_BM)
  as.numeric(t(w1)%*%B%*%F_matrix%*%t(B)+t(w1)%*%D)*2
}

Equal=function(w){
  sum(w)-1
}
EqualGradient=function(w){
  rep(1,length(w))
}

```
Initial weights and optimizer
```{r}
library(nloptr)
w0=Weight$Wgt_Port
opt=nloptr(x0=w0,eval_f=ObjFun,eval_grad_f=ObjFunGradient,
           eval_g_eq=Equal,eval_jac_g_eq=EqualGradient,
           lb=rep(0,length(w0)),ub=ifelse(w0==0,0,1),
           opts=list('algorithm'='NLOPT_LD_SLSQP','xtol_abs'=1.0e-8,'maxeval'=100000000))
w1=opt$solution
w1
```
```{r}
#w1=cbind(Weight$Ticker,w1)
#write.csv(w1,file = "F:\\Courses\\HKUST_Semester_Spring\\MAFS5210\\project2\\data_for_R\\MTE.csv",row.names = T)
```


Run the risk breakdown for the mim-TE portfolio
```{r}
TECal = function(w){
  sqrt(as.numeric(t(w)%*%B%*%F_matrix%*%t(B)%*%w+t(w)%*%D%*%w)*12)
}


ActVCal=function(w){
  sapply(Factors, function(fac){
    f=matrix(0,ncol=ncol(F_matrix),nrow=nrow(F_matrix))
    f[which(fac==Factors),]=F_matrix[which(fac==Factors),]
    as.numeric(t(w)%*%B%*%f%*%t(B)%*%w)*12
  })
}
w=as.matrix(w1-Weight$Wgt_BM)
TE=TECal(w) 
TE
ActV=ActVCal(w)
ActV
```
```{r}
u_actMTE=t(B)%*%w
u_MTE=t(B)%*%w1
u_MinTE=cbind(u_MTE,u_actMTE)
#write.csv(u_MinTE,file = "F:\\Courses\\HKUST_Semester_Spring\\MAFS5210\\project2\\data_for_R\\Data_loading_MTE.csv",row.names = T)
#write.csv(ActV,file = "F:\\Courses\\HKUST_Semester_Spring\\MAFS5210\\project2\\data_for_R\\ActV_MTE.csv",row.names = T)
```

##Step13:
MVO portfolio

We use the composite score as the alpha for stock
```{r}
Alpha=data.table(read.csv('F:\\Courses\\HKUST_Semester_Spring\\MAFS5210\\project2\\data_for_R\\Score.csv',header = T))
Alpha

setkey(Alpha,TICKER)
setkey(Weight,Ticker)
Alpha=Alpha[Weight]
Alpha[is.na(Score),Score:=0]
```

Optimization setting
```{r}
ObjFun = function(w){
  -sum(Alpha$Score*w)
}

ObjFunGradient = function(w){
  -Alpha$Score
}
Equal=function(w){
  sum(w)-1
}
EqualGradient = function(w){
  rep(1,length(w))
}
Inequal=function(w){
  w1=as.matrix(w-Weight$Wgt_BM)
  as.numeric(t(w1)%*%B%*%F_matrix%*%t(B)%*%w1+t(w1)%*%D%*%w1)-0.01/12
}
InequalGradient=function(w){
  w1=as.matrix(w-Weight$Wgt_BM)
  as.numeric(t(w1)%*%B%*%F_matrix%*%t(B)+t(w1)%*%D)*2
}
w0=Weight$Wgt_Port
opt=nloptr(x0=w0,eval_f = ObjFun,eval_grad_f = ObjFunGradient,
           eval_g_eq = Equal, eval_jac_g_eq = EqualGradient,
           eval_g_ineq = Inequal,eval_jac_g_ineq = InequalGradient,
           lb=rep(0,length(w0)),ub=ifelse(w0==0,0,1),
           opts=list('algorithm'='NLOPT_LD_SLSQP','xtol_abs'=1.0e-8,'maxeval'=100000000))
w1=opt$solution
w1

```

```{r}
ww=cbind(Weight[,Ticker],w1)
ww
u_MVO1=t(B)%*%w1
w=w1-Weight$Wgt_BM
u_actMVO=t(B)%*%w
u_MVO=cbind(u_MVO1,u_actMVO)
ActV_MVO=ActVCal(w)
#write.csv(u_MVO,file = "F:\\Courses\\HKUST_Semester_Spring\\MAFS5210\\project2\\data_for_R\\Data_loading_MVO.csv",row.names = T)
#write.csv(ActV_MVO,file = "F:\\Courses\\HKUST_Semester_Spring\\MAFS5210\\project2\\data_for_R\\ActV_MVO.csv",row.names = T)
#write.csv(ww,file = "F:\\Courses\\HKUST_Semester_Spring\\MAFS5210\\project2\\data_for_R\\MVO_weight.csv",row.names = T)
```

```{r}
#write.csv(w1,file = "F:\\Courses\\HKUST_Semester_Spring\\MAFS5210\\project2\\data_for_R\\MVO.csv",row.names = F)
```





